\paperTitle

### 2014-05-17 doc/icml2014/directed.tex

%In this section, we will develop a consistent parameter estimate for a class of directed graphical models.
%The conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
%recovered from {\TensorFactorize}, are not the underlying parameters.
  %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
  %\given h_j)$ for some $v, j$, we present a systematic approach to learn the
  %conditional probability tables for every clique for a class of directed
  %graphical models.
Having recovered conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$,
we now seek to compute the marginal distribution of sets of hidden variables
$Z_S \eqdef \Pr(\bh_S)$.
%taking us one step closer to the actual model parameters.

@@

We now show that if a subset of hidden variables $S$ has the exclusive views property,
then we can recover the marginal distribution $\Pr(\bh_S)$.
Consider any $S = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let
  $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $S$ and define $\sV
  = \{v_1, \ldots, v_m\}$. % be a set of exclusive views for the clique $S$.
By the exclusive views property,
the marginal over the observed variables $\Pr(\bx_\sV)$
factorizes according to the marginal over the hidden variables $\Pr(\bh_S)$
times the conditional moments:
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  &= \sum_{\vh_S} \Pr(\Sh{S}) 
                    \Pr(x_{v_1} \given h_{i_1}) \cdots \Pr(x_{v_m} \given h_{i_m}) \\
  &= Z_{S}(\mOpp{v_1}{i_1},\dots,\mOpp{v_m}{i_m}) \\
  &= Z_{S}(\mOppAll),
\end{align*}
where $\mOppAll = \mOpp{v_1}{i_1} \otimes \cdots \otimes \mOpp{v_m}{i_m}$ is the tensor product of
all the conditional moments.
Vectorizing, we have that
$Z_S \in \Re^{k^m}$,
$M_\sV \in \Re^{d^m}$,
and $\mOppAll \in \Re^{d^m \times k^m}$.
Since each $\mOpp{v}{i}$ has full column rank $k$,
the tensor product $\mOppAll$ has full column rank $k^m$.
Succinctly, $\mO_\sV$ (which can be estimated directly from data)
is a linear function of $Z_S$ (what we seek to recover).
We can solve for the hidden marginals $Z_S$ simply by projecting $\mO_\sV$ onto the column
space of $\mOppAll$:
\begin{align*}
  Z_{S} &= \mO_\sV(\mOpp{v_1}{i_1}\pinv,\cdots,\mOpp{v_m}{i_m}\pinv).
\end{align*}

@@

Our method extends naturally to the case in which the observed variables
  are real-valued $x_v \in \Re^d$, as long as the hidden variables remain discrete. 
In this setting, the conditional moments
  $\mOpp{v}{i} \eqdef \E(x_v \given h_i) \in \Re^{d \times k}$
  would no longer be distributions but general rank $k$ matrices.
  %for each observed variable $x_v$ and
  %hidden variable $h_i$.
%\algorithmref{learnMarginals} % and \algorithmref{directed} 
%still applies and allow us to recover clique marginals $Z_\sC$.
%  \todo{make this a footnote if running out of space}
% PL: don't understand this comment
%In this case, we only require that the hidden variables in distinct
  %cliques share parameters.

>>>>

%In this section, we will develop a consistent parameter estimate for a class of directed graphical models.
%The conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
%recovered from {\TensorFactorize}, are not the underlying parameters.
  %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
  %| h_j)$ for some $v, j$, we present a systematic approach to learn the
  %conditional probability tables for every clique for a class of directed
  %graphical models.
Having recovered conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$,
we now seek to compute the marginal distribution of sets of hidden variables
$Z_S \eqdef \Pr(\bh_S)$.
%taking us one step closer to the actual model parameters.

@@

We now show that if a subset of hidden variables $S$ has the exclusive views property,
then we can recover the marginal distribution $\Pr(\bh_S)$.
Consider any $S = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let
  $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $S$ and define $\sV
  = \{v_1, \ldots, v_m\}$. % be a set of exclusive views for the clique $S$.
By the exclusive views property,
the marginal over the observed variables $\Pr(\bx_\sV)$
factorizes according to the marginal over the hidden variables $\Pr(\bh_S)$
times the conditional moments:
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  &= \sum_{\vh_S} \Pr(\Sh{S}) 
                    \Pr(x_{v_1} | h_{i_1}) \cdots \Pr(x_{v_m} | h_{i_m}) \\
  &= Z_{S}(\mOpp{v_1}{i_1},\dots,\mOpp{v_m}{i_m}) \\
  &= Z_{S}(\mOppAll),
\end{align*}
where $\mOppAll = \mOpp{v_1}{i_1} \otimes \cdots \otimes \mOpp{v_m}{i_m}$ is the tensor product of
all the conditional moments.
Vectorizing, we have that
$Z_S \in \Re^{k^m}$,
$M_\sV \in \Re^{d^m}$,
and $\mOppAll \in \Re^{d^m \times k^m}$.
Since each $\mOpp{v}{i}$ has full column rank $k$,
the tensor product $\mOppAll$ has full column rank $k^m$.
Succinctly, $\mO_\sV$ (which can be estimated directly from data)
is a linear function of $Z_S$ (what we seek to recover).
We can solve for the hidden marginals $Z_S$ simply by projecting $\mO_\sV$ onto the column
space of $\mOppAll$:
\begin{align*}
  Z_{S} &= \mO_\sV(\mOpp{v_1}{i_1}\pinv,\cdots,\mOpp{v_m}{i_m}\pinv).
\end{align*}

@@

Our method extends naturally to the case in which the observed variables
  are real-valued $x_v \in \Re^d$, as long as the hidden variables remain discrete. 
In this setting, the conditional moments
  $\mOpp{v}{i} \eqdef \E(x_v | h_i) \in \Re^{d \times k}$
  would no longer be distributions but general rank $k$ matrices.
  %for each observed variable $x_v$ and
  %hidden variable $h_i$.
%\algorithmref{learnMarginals} % and \algorithmref{directed} 
%still applies and allow us to recover clique marginals $Z_\sC$.
%  \todo{make this a footnote if running out of space}
% PL: don't understand this comment
%In this case, we only require that the hidden variables in distinct
  %cliques share parameters.
